---
title: "Main Study Analysis: The Compatibility Effect between Physical Stimulus Size and Left-Right Responses"
author: "Florian Pätzold 977687, Daniel Menzel 979379, Ramon Zacharias 977316, Shari Hiltner 979694"
date: "24/07/2020"
output: html_document
---

This analysis script is largely based on the internet tutorial, which you can find here: [https://www.scribbr.com/statistics/anova-in-r/].

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = F, message = F, warning = F)
```

## 0. Setting up the required packages

```{r libraries}

# package for convenience functions (e.g. ggplot2, dplyr, etc.)
library(tidyverse)

library(ggplot2)
library(ggpubr)
library(broom)

# package for AIC evaluation
library(AICcmodavg)

# package to visualize 
library(bayesplot)

# global color scheme / non-optimized
project_colors = c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#000000")

# setting theme colors globally
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = project_colors)
}
scale_fill_discrete <- function(...) {
   scale_fill_manual(..., values = project_colors)
} 

# define scale function to avoid matrix object in tibble
my_scale <- function(x) c(scale(x))

```

# 1. Load in the data.

We here laod in the raw data by our 3 participants gathered in the pilot study of our experiment (and bind them together to one data frame).

```{r data}
# Load in the data from the pilot study
comp_raw <- read_csv("/Users/florian/Documents/Studium/Semester_4/Experimental Psychology Lab/Final_Project/data/compatibility_raw.csv")

# Let's have a look at the raw data set (the "key1" values should not be logical "FALSE" but character data type: "f")
comp_raw
```

# 2. Processing and cleaning the data

Time to clean our data! Find the justification of the cleaning process in the preregistration of our study.

```{r cleaning}
# We exclude columns, that are not required for the analysis
comp_data <- comp_raw %>%
  dplyr::select(submission_id,
         trial_type,
         key_pressed,
         Mapping,
         correctness,
         RT,
         handedness, # discard data from left-handers because of H2
         timeSpent # discard data from participants, that finished experiment faster than 5min
         ) %>%
  rename("response" = "key_pressed") %>% 
  rename("mapping" = "Mapping") %>%
  rename("time_spent" = "timeSpent") %>%
  filter(handedness == "right" & RT > 100 & RT < 1500 & time_spent > 5 & trial_type == "main" & correctness == "correct") # data exclusion (see preregistration)

# Change key name to response location
for (i in 1:length(comp_data$response)) {
  if(comp_data$response[i] == "j") {
    comp_data$response[i] <- "right"
  } else {
    comp_data$response[i] <- "left"
  }
}

comp_data
```

# 3. Visualizing the data.

To get a good idea of what our data looks like we now plot the mean reaction times for the different mapping conditions (compatible and incompatible) grouped by response locations (left and right).

```{r visualization}
# Define a function for calculating the standard error
se <- function(x) sd(x) / sqrt(length(x))

# aggregate mean reaction times per subject and mapping condition
comp_data_agg <- comp_data %>%
  group_by(response, mapping) %>% 
  summarise(mean_RT = mean(RT), se_RT = se(RT))

# the aggregated data frame
comp_data_agg

comp_data_agg %>% ggplot(aes(x = mapping, y = mean_RT, fill = mapping)) +
  geom_errorbar(aes(ymin = mean_RT - se_RT, ymax = mean_RT + se_RT), width = 0.2) +
  geom_col() +
  ylab("Reaction time (ms) \n") +
  xlab("\n Response location") +
  facet_grid(~response) +
  coord_cartesian(ylim = c(475,550)) +
  legend_none()
```

# 4. Analyzing the data.

We use a two-way ANOVA with factors S-R mapping (compatible, incompatible) and response (key location: left, right) as well as the interaction between both factors to analyze our data.

```{r analysis}
# specify ANOVA models
one_way <- aov(RT ~ mapping, data = comp_data)
summary(one_way)
two_way <- aov(RT ~ mapping + response, data = comp_data)
summary(two_way)
interaction <- aov(RT ~ mapping * response, data = comp_data)
summary(interaction)

# AIC evaluation
model_set <- list(one_way, two_way, interaction)
model_names <- c("one-way", "two-way", "interaction")

aictab(model_set, modnames = model_names)

#Calculate mean and standard deviation for compatible and incompatible trials
sd_RT_comp <- comp_data %>% filter(mapping == "compatible")
mean(sd_RT_comp$RT)
sd(sd_RT_comp$RT)
sd_RT_incomp <- comp_data %>% filter(mapping == "incompatible")
mean(sd_RT_incomp$RT)
sd(sd_RT_incomp$RT)
```
AIC values are the lowest for the one-way ANOVA model, but the AIC differences between the models are marginal and as we want to investigate not only the first hypothesis but also the second hypothesis we need the interaction between factors S-R Mapping and response (which is why we nevertheless choose the interaction model).


Let's check normality assumption (histogram plot), independence of observations and homogeneity assumption.

```{r analysis_assumptions}
# 1. normality assumption: plot histogram of log-transformed RTs
hist(log(comp_data$RT))

# 2. our experimental design ensures external as well as internal validity (see experimental design plan), by using counterbalancing across our "mapping condition" independent variable and within-randomisation for the independent "response" variable.

# 3. we check for homoscedasticity
par(mfrow=c(2,2))
plot(interaction)
par(mfrow=c(1,1))
```

The log-transformed RTs have the bell-shaped curve of a normal distribution, the residual means center around 0 for all plots (1 for the logarithmic scale-location plot) and they fit the Q-Q regression line neatly. Nevertheless, for larger residual values there are large deviations from the perfectly heteroscedastic regression line but as the other plots fit the assumption of heteroscedasticity, this has not too much impact.

Now, we perform Tukey's Honestly Significant Difference (TukeyHSD) post-hoc test to investigate the direction of our results.

```{r analysis_tukeyHSD}
THSD <- TukeyHSD(interaction)
THSD
```

# 5. Results

The ANOVA revealed marginal significance for the main effect of the S-R mapping conditions compatible (mean = 528 ms, sd = 196) and incompatible(mean = 514 ms, sd = 164), F value = 4.96 with a p-value of approx. 0.026 and MSE = 161486. But no significant effect of the response location could be found (p-value of approx. 0.48).
The Tukey post-hoc test reported a difference of approx. -13.9 ms CI[-26.1,-1.7] between incompatible and compatible trials. Analysis of the interaction between mapping and response location revealed a stronger interaction effect for left-hand responses although not significant (p-value = 0.31), F value = 1.03 and MSE = 33547. Incompatibility of response location and stimulus size shortened reaction times by approx. 20.1 ms, compared to the overall mean, for the left hand CI[-42.8, 2.5] and by approx. 7.5 ms for the right hand CI[-30.2,15.2]. Thus the incompatibility reduced reaction times by approx. 12.6 ms more when giving the response with the left hand.




