---
title: "Pilot Study Analysis: The Compatibility Effect between Physical Stimulus Size and Left-Right Responses"
author: "Florian Pätzold 977687, Daniel Menzel 979379, Ramon Zacharias 977316, Shari Hiltner 979694"
date: "24/07/2020"
output: html_document
---

This analysis script is largely based on the internet tutorial, which you can find here: [https://www.scribbr.com/statistics/anova-in-r/].

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = F, message = F, warning = F)
```

## 0. Setting up the required packages

```{r libraries, include=FALSE, message=FALSE, warning=FALSE}

# package for convenience functions (e.g. ggplot2, dplyr, etc.)
library(tidyverse)

library(ggplot2)
library(ggpubr)
library(broom)
library(AICcmodavg)

# package for visualization
#library(tidybayes)

# package to visualize 
library(bayesplot)

# package to extract HDIs
#library(HDInterval)

# these options help Stan run faster
#options(mc.cores = parallel::detectCores())

#devtools::install_github("michael-franke/aida-package")
#library(aida)

# package for bayesian ANOVA
#library(bayesanova)

# for use with bayesanova
#library(MCMCpack)

# use the aida-theme for plotting
#theme_set(theme_aida())

# global color scheme / non-optimized
project_colors = c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#000000")

# setting theme colors globally
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = project_colors)
}
scale_fill_discrete <- function(...) {
   scale_fill_manual(..., values = project_colors)
} 

# define scale function to avoid matrix object in tibble
my_scale <- function(x) c(scale(x))

```

# 1. Load in the data.

We here laod in the raw data by our 3 participants gathered in the pilot study of our experiment (and bind them together to one data frame).

```{r}
# Load in the data from the pilot study
comp_pilot_1_raw <- read_csv("/Users/florian/Documents/Studium/Semester_4/Experimental Psychology Lab/Final_Project/data/pilot_1.csv")
comp_pilot_2_raw <- read_csv("/Users/florian/Documents/Studium/Semester_4/Experimental Psychology Lab/Final_Project/data/pilot_2.csv")
comp_pilot_3_raw <- read_csv("/Users/florian/Documents/Studium/Semester_4/Experimental Psychology Lab/Final_Project/data/pilot_3.csv")

# Add participant ID
comp_pilot_1_raw <- comp_pilot_1_raw %>% mutate(subject_id = rep(c(1), length(comp_pilot_1_raw$RT)))
comp_pilot_2_raw <- comp_pilot_2_raw %>% mutate(subject_id = rep(c(2), length(comp_pilot_2_raw$RT)))
comp_pilot_3_raw <- comp_pilot_3_raw %>% mutate(subject_id = rep(c(3), length(comp_pilot_3_raw$RT)))

# Bind data together
comp_pilot_raw <- rbind(comp_pilot_1_raw, comp_pilot_2_raw, comp_pilot_3_raw)
comp_pilot_raw
```

# 2. Processing and cleaning the data

Time to clean our data! Find the justification of the cleaning process in the preregistration of our study.

```{r}
# We exclude columns, that are not required for the analysis
comp_pilot_data <- comp_pilot_raw %>%
  dplyr::select(subject_id,
         trial_type,
         key_pressed,
         Mapping,
         correctness,
         RT,
         handedness, # discard data from left-handers because of H2
         timeSpent # discard data from participants, that finished experiment faster than 5min
         ) %>%
  mutate(RT = RT - 2500) %>%  # We forgot to subtract pause (1500ms) and fix_duration (1000ms) hooks in the magpie implementation of the experiment.
  rename("response" = "key_pressed") %>% 
  rename("mapping" = "Mapping") %>%
  rename("time_spent" = "timeSpent") %>%
  filter(handedness == "right" & RT > 100 & RT < 1500 & time_spent > 5 & trial_type == "main" & correctness == "correct") # data exclusion (see preregistration)

# Change key name to response location
for (i in 1:length(comp_pilot_data$response)) {
  if(comp_pilot_data$response[i] == "j") {
    comp_pilot_data$response[i] <- "right"
  } else {
    comp_pilot_data$response[i] <- "left"
  }
}

comp_pilot_data
```

# 3. Visualizing the data.

To get a good idea of what our data looks like we now plot the mean reaction times for the different mapping conditions (compatible and incompatible) grouped by response locations (left and right).

```{r}
# Define a function for calculating the standard error
se <- function(x) sd(x) / sqrt(length(x))

# aggregate mean reaction times per subject and mapping condition
comp_pilot_agg <- comp_pilot_data %>%
  group_by(response, mapping) %>% 
  summarise(mean_RT = mean(RT), se_RT = se(RT))

# the aggregated data frame
comp_pilot_agg

comp_pilot_agg %>% ggplot(aes(x = mapping, y = mean_RT, fill = mapping)) +
  geom_errorbar(aes(ymin = mean_RT - se_RT, ymax = mean_RT + se_RT), width = 0.2) +
  geom_col() +
  ylab("Reaction time (ms) \n") +
  xlab("\n Response location") +
  facet_grid(~response) +
  coord_cartesian(ylim = c(450,550)) +
  legend_none()
```

# 4. Analyzing the data.

We use a two-way ANOVA with factors S-R mapping (compatible, incompatible) and response (key location: left, right) as well as the interaction between both factors to analyze our data.

```{r}
# specify ANOVA models
one_way <- aov(RT ~ mapping, data = comp_pilot_data)
summary(one_way)
two_way <- aov(RT ~ mapping + response, data = comp_pilot_data)
summary(two_way)
interaction <- aov(RT ~ mapping * response, data = comp_pilot_data)
summary(interaction)

# AIC evaluation
model_set <- list(one_way, two_way, interaction)
model_names <- c("one-way", "two-way", "interaction")

aictab(model_set, modnames = model_names)

#Calculate mean and standard deviation for compatible and incompatible trials
sd_RT_comp <- comp_pilot_data %>% filter(mapping == "compatible")
mean(sd_RT_comp$RT)
sd(sd_RT_comp$RT)
sd_RT_incomp <- comp_pilot_data %>% filter(mapping == "incompatible")
mean(sd_RT_incomp$RT)
sd(sd_RT_incomp$RT)
```
AIC values are the lowest for the one-way ANOVA model, but as we want to investigate not only the first hypothesis but also the second hypothesis we need the interaction between factors S-R Mapping and response (which is why we nevertheless choose the interaction model).


Let's check normality assumption (histogram plot), independence of observations and homogeneity assumption.

```{r}
# 1. normality assumption: plot histogram of log-transformed RTs
hist(log(comp_pilot_data$RT))

# 2. our experimental design ensures external as well as internal validity (see experimental design plan), by using counterbalancing across our "mapping condition" independent variable and within-randomisation for the independent "response" variable.

# 3. we check for homoscedasticity
par(mfrow=c(2,2))
plot(interaction)
par(mfrow=c(1,1))
```

The log-transformed RTs have the bell-shaped curve of a normal distribution, the residual means center around 0 for all plots (1 for the logarithmic scale-location plot) and they fit the Q-Q regression line neatly.

Now, we perform Tukey's Honestly Significant Difference (TukeyHSD) post-hoc test to investigate the direction of our results.

```{r}
THSD <- TukeyHSD(interaction)
THSD
```

# 5. Results

The ANOVA revealed marginal significance for the main effect of the S-R mapping conditions compatible(mean = 473 ms, sd = 96) and incompatible(mean = 504 ms, sd = 127), F value = 6.4 with a p-value of approx. 0.012 and MSE = 81438. But no significant effect of the response location could be found. The Tukey post-hoc test reported a difference of approx. 30.7 ms CI[6.9,54.3] between incompatible and compatible trials. Analysis of the interaction between mapping and response location revealed a stronger interaction effect for left-hand responses although not significant (p-value = 0.504), F value = 0.447 and MSE = 5637. Compatibility of response location and stimulus size shortened reaction times by approx 38.4 ms, compared to the overall mean, for the left hand CI[-5.4, 83] and by approx. 22.7 ms for the right hand CI[-21.1,66.6]. Thus the compatibility reduced reaction times by approx. 16.1 ms more when giving the response with the right hand.




